Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19

### 2023-05-05 manuscript/appendix.qmd
Add human rights models summaries + tables + inline stats

tar_load(c(policies_table_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))

@@

```{r tbl-policies-summary}
#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries
#| include: true
#| echo: false
tbl_part1 <- policies_table_data$policies_pred_tbl %>% 
  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, time, derogation_ineffect, num) %>% 
  pivot_wider(names_from = "time", values_from = "num")

@@

>>>>

tar_load(c(policies_table_data, hr_table_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))

@@

```{r tbl-policies-summary}
#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for emergency policy models
#| include: true
#| echo: false
tbl_part1 <- policies_table_data$policies_pred_tbl %>% 
  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, time, derogation_ineffect, num) %>% 
  pivot_wider(names_from = "time", values_from = "num")

@@

```{r tbl-hr-summary}
#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for human rights models
#| include: true
#| echo: false
tbl_part1 <- hr_table_data$hr_pred_tbl %>% 
  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, time, group = category, derogation_ineffect, num) %>% 
  pivot_wider(names_from = "time", values_from = "num") %>% 
  mutate(group = recode(group, "NA" = "—"))

tbl_part2 <- hr_table_data$hr_mfx_tbl %>% 
  mutate(across(c(draw, min, max, p_gt_0), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(group = recode(group, "Logit" = "—")) %>% 
  mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, year_week_day, group, diff_type, num, p_gt_0) %>% 
  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_week_day")) %>% 
  mutate(derogation_ineffect = "No")

tbl_part1 %>% 
  left_join(tbl_part2, by = join_by(nice, derogation_ineffect, group)) %>% 
  mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
  select(derogation_ineffect, group, `March 2020`, `June 2021`,
         `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
         `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`) %>% 
  kbl(col.names = c("Derogation", "Level", "March 2020", "June 2021",
                    "Week", "Difference", "p > 0", "Week", "Difference", "p > 0"),
      align = "lcccccccc",
      booktabs = TRUE, linesep = "",
      escape = FALSE) %>% 
  kable_styling(font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
  column_spec(4, border_right = T) %>% 
  pack_rows(index = table(fct_inorder(tbl_part1$nice))) %>%
  add_header_above(c(" " = 2, "Predicted probabilities" = 2, 
                     "Smallest difference" = 3, "Largest difference" = 3))
```

### 2023-05-05 manuscript/manuscript.qmd
oops load this object
Add human rights models summaries + tables + inline stats

tar_load(c(policies_plot_data, human_rights_plot_data))
tar_load(c(derogation_map_data, derogation_count))

@@

policies_mfx <- policies_table_data$policies_mfx_tbl %>% 
  split(~outcome + rank)
```

@@

The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-human-rights). There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement. Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an 85% predicted probability of no discriminatory policy in March 2020 rising to nearly 100% by June 2021. The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status. Moreover, countries that derogate were substantially less likely to engage in no abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic. Minor and moderate levels of excessive enforcement are each several percentage points more likely for derogating countries than for their non-derogating counterparts.

@@

In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status. For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability). The middle ground of only minor issues with time limits is exceptionally unlikely. Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues. The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish. Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights ($p(\Delta < 0) \approx 0.83$)

>>>>

tar_load(c(policies_plot_data, human_rights_plot_data))
tar_load(c(policies_table_data, hr_table_data))
tar_load(c(derogation_map_data, derogation_count))

@@

hr_pred <- hr_table_data$hr_pred_tbl %>% 
  split(~outcome + derogation_ineffect + category + rank, drop = TRUE)

policies_mfx <- policies_table_data$policies_mfx_tbl %>% 
  split(~outcome + rank)

hr_mfx <- hr_table_data$hr_mfx_tbl %>% 
  split(~outcome + group + rank, drop = TRUE)
```

@@

The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-human-rights). There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement. Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an `r hr_pred$discriminatory_policy.No.None.min$draw_nice`% predicted probability of no discriminatory policy in March 2020 rising to `r hr_pred$discriminatory_policy.No.None.max$draw_nice`% by June 2021. The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status. Moreover, countries that derogate were substantially less likely to engage in no abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic. Minor and moderate levels of excessive enforcement are each several percentage points more likely for derogating countries than for their non-derogating counterparts.

@@

In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status. For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a `r hr_pred$no_time_limit_measures.No.None.max$draw_nice`% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with `r hr_pred$no_time_limit_measures.No.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.No.Moderate.max$draw_nice`% probability). The middle ground of only minor issues with time limits is exceptionally unlikely. Countries that formally derogated, on the other hand, have a `r hr_pred$no_time_limit_measures.Yes.None.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.None.max$draw_nice`% probability of having no issues with time limits throughout the 15 months, and only an `r hr_pred$no_time_limit_measures.Yes.Moderate.min$draw_nice`–`r hr_pred$no_time_limit_measures.Yes.Moderate.max$draw_nice`% probability of moderate issues. The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish. Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights (`r hr_mfx$non_derogable_rights.Logit.max$p_lt`)
